---
title: "Classification of P. falciparum infection recurrences with MalReBay"
author: "Monica Golumbeanu, Veronica Adhiambo Ochieng"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    toc-location: left
    number-sections: true
    code-fold: false
    code-tools: true
    code-line-numbers: true
    theme: cosmo
    highlight-style: github
    df-print: paged
    fig-width: 9
    fig-height: 5
execute:
  warning: false
  message: false
---

## Overview

**MalReBay** is an R package for classifying recurrent *Plasmodium falciparum* infections in Therapeutic Efficacy Studies (TES). For each patient, it estimates the **posterior probability that a recurrence is a recrudescence** (treatment failure) versus a new infection (reinfection).

MalReBay uses a fully Bayesian algorithm solved with Markov Chain Monte Carlo (MCMC) sampling. Key advantages:

-   Handles **polyclonal infections** (multiple parasite strains per patient)
-   Imputes **missing genotypes** as part of the MCMC
-   Supports two data types:
    -   **Length-polymorphic markers**: microsatellites, MSP1, MSP2, GLURP (allele sizes as numbers)
    -   **Amplicon sequencing (AmpSeq)**: discrete haplotypes

This notebook walks through a complete analysis of MSP/GLURP length-polymorphic data using the example dataset bundled with the package.

------------------------------------------------------------------------

## Installation

CmdStan must be installed before MalReBay, since MalReBay's Stan model is compiled during its own installation (see the package README for details).

```{r}
#| eval: false
# 1. Install CmdStan first (one-time, ~5 minutes)
cmdstanr::install_cmdstan()

# 2. Install MalReBay from GitHub (development version)
remotes::install_github("SwissTPH/MalReBay")
```

```{r}
library(MalReBay)
```

------------------------------------------------------------------------

## Input data

MalReBay requires up to three inputs: a **genotyping data file**, a **marker metadata file**, and an **MCMC configuration**. All three can be Excel (`.xlsx`) files or, for the MCMC configuration, an R named list.

### Genotyping data file

The main input is an Excel workbook. The first sheet must contain one row per sample with the following column layout:

| Column 1  | Column 2 | Column 3 | Columns4+                                  |
|-----------|----------|----------|--------------------------------------------|
| PatientID | Day      | Site     | Allele columns (`marker_1`, `marker_2`, …) |

Each patient must have exactly two rows:

-   A **Day 0** row
-   A **recurrence** row

**Allele column naming.** Columns are named `<marker_id>_1`, `<marker_id>_2`, … up to the maximum multiplicity of infection (MOI). For example, three columns `msp1_1`, `msp1_2`, `msp1_3` encode up to three distinct MSP1 alleles per sample.

-   For **length-polymorphic** data, values are numeric allele sizes (e.g. `200`, `230.5`).
-   For **AmpSeq** data, values are haplotype strings (e.g. `"Haplotyp1"`, etc.).

Missing values are recognised and set to `NA` automatically for: `N/A`, `-`, `NA`, `na`, `""`, `" "`.

A second sheet (optional, for length-polymorphic markers only) can supply additional Day 0 samples.

```{r}
# Preview the example MSP/GLURP dataset
example_file   <- "Angola_2021_TES_7NMS.xlsx"
marker_file    <- "makers_details.xlsx"

imported_data <- import_data(
  filepath        = example_file,
  marker_filepath = marker_file,
  verbose         = TRUE
)

# The imported object is a list with four elements
names(imported_data)
```

```{r}
# Main data: one row per sample (Day 0 and recurrence)
head(imported_data$late_failures)
```

```{r}
# Detected data type
imported_data$data_type
```

### Marker metadata file

The marker metadata file is an Excel file with one row per marker. Required columns:

| Column | Description |
|------------------------------------|------------------------------------|
| `marker_id` | Marker name: must match the prefix of the allele columns in the genotyping file |
| `marker_type` | Marker type: microsatellite,msp_glurp or ampseq |
| `repeatlength` | Repeat unit length (bp) for microsatellites; gap threshold for MSP/GLURP |
| `binning_method` | `"microsatellite"`, `"msp_glurp"`, or `"exact"` for each data type, respectively |

```{r}
# Inspect marker metadata
print(imported_data$marker_info)
```

### MCMC configuration

The MCMC behaviour is controlled by a set of parameters, supplied either as an Excel file or a named R list.

```{r}
# Read and display the default configuration
config_file <- "default_mcmc_config.xlsx"
cfg <- as.data.frame(readxl::read_excel(config_file))
cfg
```

| Parameter | Description |
|------------------------------------|------------------------------------|
| `n_chains` | Number of independent MCMC chains. More chains improve convergence diagnostics. |
| `chunk_size` | Number of MCMC iterations run per convergence-check cycle. |
| `max_iterations` | Hard upper limit on total MCMC iterations per site. |
| `burn_in_frac` | Fraction of samples discarded as burn-in (e.g. `0.5` → first 50 % discarded). |
| `rhat_threshold` | Gelman–Rubin R-hat upper bound for convergence (typically `1.05`). |
| `ess_threshold` | Minimum bulk effective sample size required for convergence. |
| `thinning_interval` | how many samples are skipped during thinning |
| `record_hidden_alleles` | If `TRUE`, saves the imputed latent allele states for every MCMC sample. Increases memory use. |

**Convergence rule.** The algorithm runs in chunks of `chunk_size` iterations. After each chunk it checks three criteria simultaneously: rank-normalised R-hat, bulk ESS, and tail ESS. If all pass, sampling stops early. As a fallback it also checks classical Gelman–Rubin and Geweke Z-scores. If none of the criteria are met by `max_iterations`, the run completes with a warning.

You can also pass the configuration directly as a named list instead of an Excel file.

------------------------------------------------------------------------

## Running the pipeline

The main package function `MalReBay()` runs data import, MCMC classification, result summarisation, and output saving.

```{r}
quick_config <- list(
  n_chains     = 2,
  iter         = 1000,
  burn_in_frac = 0.5,
  adapt_delta  = 0.8,
  random_seed  = 42
)

results <- MalReBay(
  filepath        = example_file,
  marker_filepath = marker_file,
  mcmc_config     = quick_config,
  output_folder   = "bayesian_algorithm_output",
  verbose         = TRUE
)
```

### Parameters

| Parameter | Description |
|------------------------------------|------------------------------------|
| `filepath` | Path to the genotyping Excel file. |
| `marker_filepath` | Path to the marker metadata Excel file. |
| `mcmc_config` | Path to the MCMC configuration Excel file, or a named list. |
| `output_folder` | Directory where all output files and plots are written. Set to `NULL` to suppress file saving. |
| `n_workers` | Number of parallel workers for multi-chain MCMC. Capped at the number of physical cores. |
| `verbose` | If `TRUE`, prints progress and convergence messages to the console. |

------------------------------------------------------------------------

## Output

`MalReBay()` returns a named list with four elements and, when `output_folder` is set, saves the CSV files and diagnostic plots.

### Return value

```{r}
names(results)
```

#### `posterior_probabilities`

One row per patient. The key column is `Probability`: the posterior probability that the recurrence is a **recrudescence**. Also includes per-patient locus availability counts.

```{r}
head(results$posterior_probabilities)
```

A probability close to **1** indicates recrudescence; close to **0** indicates reinfection. Values near **0.5** reflect uncertainty, often due to limited or missing genotyping data.

#### `comparison`

Combines the Bayesian posterior probabilities with the traditional **allele match counting** result for each patient, enabling side-by-side comparison of both methods.

```{r}
head(results$comparison)
```

#### `convergence`

Per-site MCMC convergence diagnostics: Gelman–Rubin R-hat and effective sample size.

```{r}
results$convergence
```

An R-hat below `rhat_threshold` (typically 1.05) and ESS above `ess_threshold` indicate that the chains have mixed well and the posterior estimates are reliable.

#### `mcmc_loglikelihoods`

Raw per-chain log-likelihood traces, one list element per site. Useful for manual convergence inspection.

```{r}
# Sites present in the data
names(results$mcmc_loglikelihoods)
```

### Output files

When `output_folder` is provided, the following files are written:

| File | Contents |
|------------------------------------|------------------------------------|
| `posterior_probabilities.csv` | Posterior probability of recrudescence per patient, with locus counts. |
| `bayesian_match_counting_comparison.csv` | Bayesian results alongside traditional match counting. |
| `mcmc_convergence_summary.csv` | R-hat and ESS per site. |
| `recrudescence_probability_histogram.png` | Distribution of posterior probabilities across all patients. |
| `diversity_length_polymorphic_comparison.png` | Allele/haplotype diversity pie charts per marker. |
| `moi_per_marker_by_site.png` | Violin plots of multiplicity of infection per marker and site. |
| `convergence_diagnosis/<site>/` | Per-site MCMC trace plots, Gelman–Rubin plot, log-likelihood histogram, and autocorrelation plots. |

### Plots

The pipeline automatically generates and displays the following plots.

**Allele diversity**: pie charts showing the frequency distribution of each allele (or haplotype) across all samples:

```{r}
p_div <- plot_markers_diversity(
  genotypedata  = imported_data$late_failures,
  data_type     = imported_data$data_type,
  marker_info   = imported_data$marker_info,
  output_folder = NULL
)
print(p_div)
```

**Multiplicity of infection (MOI)**: violin plots showing the number of distinct alleles per marker per sample, split by site:

```{r}
p_moi <- plot_moi(imported_data$late_failures, output_folder = NULL)
print(p_moi)
```

**Posterior probability histogram**: distribution of recrudescence probabilities across all classified patients:

```{r}
p_hist <- plot_probability_histogram(results, output_folder = NULL)
```

**MCMC convergence diagnostics**: trace plots, Gelman–Rubin plots, log-likelihood histograms, and autocorrelation plots saved under `convergence_diagnosis/<site>/` when `output_folder` is set.

------------------------------------------------------------------------

## Interpreting the results

A standard classification rule applied to `Probability`:

| Probability | Classification |
|-------------|----------------|
| ≥ 0.5       | Recrudescence  |
| \< 0.5      | Reinfection    |

```{r}
pp <- results$posterior_probabilities
pp$Classification <- ifelse(pp$Probability >= 0.5, "Recrudescence", "Reinfection")
table(pp$Classification)
```

Patients with few comparable loci (`N_Comparable_Loci`) should be interpreted with caution because the posterior is driven by less information and will be closer to the prior.

```{r}
# Flag low-information classifications
pp[pp$N_Comparable_Loci <= 1, c("Sample.ID", "Site", "Probability",
                                  "N_Comparable_Loci", "Classification")]
```

If the convergence summary shows high R-hat values or low ESS for a site, consider increasing `max_iterations` or `chunk_size` in the MCMC configuration.
