---
title: "MalReBay analysis template: MSP1/MSP2/GLURP markers"
author: "Monica Golumbeanu"
format:
  html:
    toc: true
    toc-depth: 2
    code-fold: false
execute:
  warning: false
  message: false
---

::: callout-note
Lightweight starting template for an MSP1/MSP2/GLURP panel. For explanations of every step, see the [full tutorial](https://swisstph.github.io/MalReBay/articles/MalReBay.html). This example uses the example datasets provided with the MalReBay package. Replace the file paths in **Setup** with your own data.
:::

## Setup

```{r}
library(MalReBay)

# --- edit these for your own data -------------------------------------
filepath        <- system.file("extdata", "Dataset_msp_glurp.xlsx", package = "MalReBay")
marker_filepath <- system.file("extdata", "makers_details.xlsx", package = "MalReBay")
output_folder   <- "malrebay_results"
# ------------------------------------------------------------------------
```

## 1. Load data

```{r}
imported_data <- import_data(
  filepath        = filepath,
  marker_filepath = marker_filepath,
  verbose          = TRUE
)

head(imported_data$late_failures)
```

::: callout-tip
`detect_msp_variants()` recognises the MSP1 family variants (`K1`, `MAD20`, `RO33`) and MSP2 family variants (`3D7`, `FC27`, `IC`) from the marker column names automatically.
:::

## 2. Descriptive analysis

```{r}
# Plot the MOI distributions for each site
moi_plots <- plot_moi(imported_data$late_failures, output_folder = NULL)
moi_plots$A
moi_plots$B
moi_plots$C
```

```{r}
# Plot the diversity of each marker
diversity_plot <- plot_markers_diversity(
  genotypedata  = imported_data$late_failures,
  data_type     = imported_data$data_type,
  marker_info   = imported_data$marker_info,
  output_folder = NULL
)
diversity_plot
```

```{r}
#| fig-width: 13
#| fig-height: 6
#| out-width: 80%

# Plot the distributions of alleles
allele_plot <- plot_allele_distribution(
  genotypedata  = imported_data$late_failures,
  marker_info   = imported_data$marker_info,
  output_folder = NULL
)
allele_plot
```

## 3. Run infection recurrence classification

Uses the package's bundled default MCMC settings. If convergence looks poor in the next section, pass a stronger `mcmc_config` (e.g., `list(n_chains = 4, iter = 3000, adapt_delta = 0.95)`) and rerun.

**This step might take a while depending on your dataset size, number of chains and iterations.**

```{r}
# Classify recurrent infections
results <- MalReBay(
  filepath        = filepath,
  marker_filepath = marker_filepath,
  output_folder   = output_folder,
  verbose         = TRUE
)
```

## 4. Inspect convergence

```{r}
# Raw convergence metrics:
results$convergence
```

```{r}
#| fig-width: 12
#| fig-height: 8

# Plotting convergence diagnostics for the first site. Replace the site name for visualising the diagnostics from other sites.
site_name <- names(results$mcmc_loglikelihoods)[1]

plot_likelihood_diagnostics(
  all_chains_loglikelihood = results$mcmc_loglikelihoods[[site_name]],
  site_name                = site_name,
  save_plot                = FALSE,
  verbose                  = FALSE,
  combine_plots            = TRUE
)$combined
```

## 5. Visualize results

```{r}
#| fig-width: 12
#| fig-height: 8
plot_probability_histogram(results, output_folder = NULL)
```

```{r}
plot_comparison_heatmap(results, imported_data$marker_info, output_folder = NULL)
```

## 6. Explore results

```{r}
results$posterior_probabilities
```

```{r}
results$comparison
```

All tables and plots are also saved under `output_folder` (set `output_folder = NULL` above to skip saving and only view results here).
